Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-24 ms.tex
lx

The X-ray luminosities of galactic nuclei provide a powerful diagnostic of the SMBH accretion rates and how they vary across different galaxy properties and SMBH masses (e.g., \citealt{Ho08} and references therein).  Figure \ref{fig:vwSources} shows that the heating associated with the average star formation history of a galaxy is approximately constant with SMBH mass, except for the largest black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For a fixed wind heating parameter, the Eddington ratio $\dot{M}/\dot{M}_{\rm edd}$ increases $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  This dependence, shown directly in Figure XXX, appears to contradict the observed behavior of the nuclear X-ray luminosities of quiescent galaxies, which (though showing several orders of magnitude scatter) instead on average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where $\alpha \in [-0.4,-0.2]$ \citep{Zhang+09, Pellegrini10, Miller+15}.

Figure~\ref{fig:miller} shows our predictions for $\langle L_{X}\rangle =\epsilon_X \epsilon_{\rm bol} \Mdot c^2$ in Eddington units as a function of black hole mass for core (blue) and cusp (green) galaxies, where $\epsilon_{\rm bol} = L_{X}/L_{\rm bol}= 0.1$ is the assumed bolometric correction appropriate for low luminosity AGN (\citealt{Ho08}).  The left panel is calculated assuming a constant low value for the radiative efficiency $\epsilon_X = 10^{-4}$, typical of those estimated for low luminosity AGN (e.g.,
\citealt{DiMatteo+00}, \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}, \citealt{Quataert:2004a}).  The luminosity shown in the right
panel instead assumes an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$ that depends on $\dot{M}$ in the way predicted by models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}).  Superimposed with the theoretical curves
are symbols showing X-ray measurements (stars) and upper limits
(triangles) compiled by \citet{Miller+15}, along with the best power-law
fit (red line) and the X-ray detection threshold (gray line).  Finally, purple, orange, and {\bf brown} lines
orange lines show the maximum accretion rates for thermally stable
accretion (eq.~[\ref{eq:Mdotmax}]), as set by SN Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), and as allowed by Compton heating feedback (eq.~[\ref{eq:MdotC}]), respectively.

To zeroth order our predicted values of $\langle L_{X}/L_{\rm edd}\rangle$ for average star formation histories overlap with the data reasonably well.  However, our predicted increase of $\langle L_{X}/L_{\rm edd}\rangle$ with black hole mass appears to be in disagreement with observed decrease of $\langle L_{X}/L_{\rm edd}\rangle$ with $M_{\bullet}$ for both core and cusp galaxies.  Better agreement is obtained by taking into account the fact that core(cusp) galaxie are more commonly associated with larger(smaller) black hole masses, such that when both samples are combined one predicts a flatter dependence of $\langle L_X/L_{\rm edd}\rangle$ of $M_{\bullet}$.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically 2-4 orders of magnitude is also natural within a steady-accretion scenario due to the strong sensitivity of the Eddington ratio on the heating rate, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) and the likewise sensitive dependence of the heating rate on the intermittency of the star formation history (Fig.~NICK).  Alternatively, in the case of an old stellar population incapable of supporting steady accretion, resulting in cyclic accretion episodes (e.g.~\citealt{Ciotti+10}), this large scatter can be understood by the wind range of gas masses experienced over the course of a cycle.  The cyclic model also predicts that a higher fraction of $L_{X}$ nuclei with smaller stellar ages, consistent with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations host less X-ray luminous nuclei.  However, this observation can also be understood in the stable accretion picture by the fact that young stars provide sufficient stellar wind heating to produce a thermally stable accretion state.

@@

>>>>

The X-ray luminosities of galactic nuclei provide a powerful diagnostic of the SMBH accretion rates and how they vary across different galaxy properties and SMBH masses (e.g.,  \citealt{Ho08}  and references therein).  Figure \ref{fig:vwSources} shows that the heating associated with the average star formation history of a galaxy is approximately constant with SMBH mass, except for the largest black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For a fixed wind heating parameter, the Eddington ratio $\dot{M}/\dot{M}_{\rm edd}$ increases $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  This dependence, shown directly in Figure XXX, appears to contradict the observed behavior of the nuclear X-ray luminosities of quiescent galaxies, which (though showing several orders of magnitude scatter) instead on average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where $\alpha \in [-0.4,-0.2]$ \citep{Zhang+09, Pellegrini10, Gallo+10, Miller+15}.

Figure~\ref{fig:miller} shows our predictions for $\langle L_{X} \rangle =\epsilon_X \epsilon_{\rm bol} \dot{M} c^2$ in Eddington units as a function of black hole mass for core (blue) and cusp (green) galaxies, where $\epsilon_{\rm bol} = L_{X}/L_{\rm bol}= 0.1$ is the assumed bolometric correction appropriate for low luminosity AGN (\citealt{Ho08}).  The left panel is calculated assuming a constant low value for the radiative efficiency $\epsilon_X = 10^{-4}$, typical of those estimated for low luminosity AGN (e.g., \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}).  The luminosity shown in the right
panel instead assumes an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$ that depends on $\dot{M}$ in the way predicted by models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}).  Plotted next to the theoretical curves are the X-ray measurements (stars) and upper limits
(triangles) from \citet{Miller+15} (cf.~\citealt{Gallo+10}), along with the resulting best power-law
fit (red line) and the X-ray detection threshold (gray line).  Purple, orange, and {\bf brown} lines show the maximum accretion rates for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]), as set by SN Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), and as allowed by Compton heating feedback (eq.~[\ref{eq:MdotC}]), respectively.

To first order our predicted values of $\langle L_{X}/L_{\rm edd}\rangle$ for the average galactic star formation histories overlap with the data reasonably well.  Our prediction that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass appears to be in disagreement with observed decrease of $\langle L_{X}/L_{\rm edd}\rangle$ with $M_{\bullet}$ for both core and cusp galaxies.  However, better agreement is obtained by taking into account the fact that core(cusp) galaxie are more commonly associated with larger(smaller) black hole masses, such that when both samples are combined one predicts a flatter dependence of $\langle L_X/L_{\rm edd}\rangle$ of $M_{\bullet}$.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically 2-4 orders of magnitude can also be understood within a steady-accretion scenario due to the strong sensitivity of the accretion rate on the efficacy of gas heating, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) and the equally sensitive dependence of the heating rate on stochastic intermittency in the star formation history (Fig.~NICK).

In the case of an old stellar population which produces insufficient heating to support steady accretion, this large scatter can in principle also be understood by the wide range of gas masses experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  The cyclic model also predicts that high $L_{X}$ nuclei should be more common in galaxies with younger stellar ages.  This appears in tension with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations are observed to host less X-ray luminous nuclei, at least for stellar ages $<3$ Gyr.  This observation can also be understood in the stable accretion picture by the fact that young stars provide sufficient stellar wind heating to produce a thermally stable accretion state.

@@

\subsection{Steady Accretion versus Cyclic Outbursts}

GALEX observations have shown that a small fraction of early-type galaxies in the local universe have undergone recent ($< 1-2$ Gyr old) star formation episodes (\citealt{Donas+07}).
